<!-- %:::% index begin %:::% -->
# Welcome {.unnumbered}

```{r}
#| label: setup
#| include: false

source(here::here("R", "_setup.R"))
```

You are viewing the web version of this thesis. To view the PDF version, click [here](index.pdf) or use the PDF icon in the menu.

All analyses presented in this document are fully reproducible and were conducted using the [R programming language](https://www.r-project.org/) and the [Quarto](https://quarto.org/) publishing system.

To explore the code and repository for this thesis, click [here](https://github.com/danielvartan/mastersthesis) or use the GitHub icon in the menu. The research compendium is also accessible via [The Open Science Framework](https://osf.io/) by clicking [here](https://doi.org/10.17605/OSF.IO/YGKTS).

## License

[![](https://img.shields.io/badge/License-CC_BY_4.0-lightgrey.svg)](https://creativecommons.org/licenses/by/4.0/){fig-align="left"}

This work is licensed under a [Creative Commons Attribution 4.0 International License](https://creativecommons.org/licenses/by/4.0/). This means you are free to share and adapt the work for any purpose, including commercial use, provided you give appropriate credit, link to the license, and indicate any changes made.

## Citation

To cite this work, please use the following format:

Vartanian, D. V. (`r env_vars[["year"]]`). *`r env_vars[["title"]]`* [`r env_vars[["academic-title"]]`'s `r env_vars[["type-of-work"]]`, `r env_vars[["university"]]`]. [https://doi.org/10.17605/OSF.IO/YGKTS](https://doi.org/10.17605/OSF.IO/YGKTS)

BibTeX citation:

```{r}
#| include: false

bib_tag <- ifelse(
  grepl(env_vars$`academic-title`, "Master", ignore.case = TRUE), 
  "@mastersthesis", 
  "@phdthesis"
  )
```

```
`r bib_tag`{vartanian`r env_vars[["year"]]`,
  title = {`r env_vars[["title"]]`},
  author = {Daniel Vartanian},
  year = {`r env_vars[["year"]]`},
  address = {`r env_vars[["location"]]`},
  school  = {`r env_vars[["university"]]`},
  langid = {`r env_vars[["language"]]`},
  url = {https://doi.org/10.17605/OSF.IO/YGKTS},
  note = {`r env_vars[["version-note"]]`}
}
```
<!-- %:::% index end %:::% -->
