<!-- %:::% index begin %:::% -->
# Welcome {.unnumbered}

```{r}
#| label: setup
#| include: false

source(here::here("R/_setup.R"))
```

You are viewing the web version of this thesis. Click [here](index.pdf) or on the PDF icon in the menu to see the PDF version.

All analyses in this document are reproducible and were conducted using the [R programming language](https://www.r-project.org/) along with the [Quarto](https://quarto.org/) publishing system. Click [here](https://github.com/danielvartan/mastersthesis) or on the GitHub icon in the menu to access the research compendium for the thesis.

## Citation

For attribution, please cite this work as:

`r env_vars[["author-surname"]]`, `r sub("\\. .\\.$", "", env_vars[["author-initials"]])`. (`r env_vars[["year"]]`). _`r env_vars[["title"]]`_ [`r env_vars[["academic-title"]]`'s `r env_vars[["type-of-work"]]`, `r env_vars[["university"]]`]. `r env_vars[["url"]]`

BibTeX citation:

```{r}
#| include: false

bib_tag <- ifelse(
  grepl(env_vars$`academic-title`, "Master", ignore.case = TRUE), 
  "@mastersthesis", 
  "@phdthesis"
  )
```

```
`r bib_tag`{`r tolower(env_vars[["author-surname"]])``r env_vars[["year"]]`,
  title = {`r env_vars[["title"]]`},
  author = {`r env_vars[["author"]]`},
  year = {`r env_vars[["year"]]`},
  address = {`r env_vars[["location"]]`},
  school  = {`r env_vars[["university"]]`},
  langid = {`r env_vars[["language"]]`},
  url = {`r env_vars[["url"]]`},
  note = {`r env_vars[["version-note"]]`}
}
```
<!-- %:::% index end %:::% -->
