:::: {.panel-tabset}
### MSF~sc~ (Chronotype proxy) (seconds)

::: {#tbl-sm-5-var-distribution-stats-msf-sc}
```{r}
#| code-fold: true

data |>
  stats_summary(
    col = 'msf_sc',
    name = 'MSF~sc~ (Chronotype proxy) (seconds)',
    as_list = FALSE
  )
```

Statistics for the `msf_sc` variable.
:::

::: {#fig-sm-5-var-distribution-histogram-msf-sc}
```{r}
#| code-fold: true

data |>
  test_normality(
    col = 'msf_sc',
    name = 'MSF~sc~ (Chronotype proxy) (seconds)'
  )
```

Histogram of the `msf_sc` variable with a kernel density
estimate, along with a quantile-quantile (Q-Q) plot between the
variable and the theoretical quantiles of the normal distribution.
:::

::: {#fig-sm-5-var-distribution-box-plot-msf-sc}
```{r}
#| code-fold: true

data |>
  plotr:::plot_box_plot(
    col = 'msf_sc'
  )
```

Boxplot of the `msf_sc` variable.
:::

### Age (years)

::: {#tbl-sm-5-var-distribution-stats-age}
```{r}
#| code-fold: true

data |>
  stats_summary(
    col = 'age',
    name = 'Age (years)',
    as_list = FALSE
  )
```

Statistics for the `age` variable.
:::

::: {#fig-sm-5-var-distribution-histogram-age}
```{r}
#| code-fold: true

data |>
  test_normality(
    col = 'age',
    name = 'Age (years)'
  )
```

Histogram of the `age` variable with a kernel density
estimate, along with a quantile-quantile (Q-Q) plot between the
variable and the theoretical quantiles of the normal distribution.
:::

::: {#fig-sm-5-var-distribution-box-plot-age}
```{r}
#| code-fold: true

data |>
  plotr:::plot_box_plot(
    col = 'age'
  )
```

Boxplot of the `age` variable.
:::

### Longitude (decimal degrees)

::: {#tbl-sm-5-var-distribution-stats-longitude}
```{r}
#| code-fold: true

data |>
  stats_summary(
    col = 'longitude',
    name = 'Longitude (decimal degrees)',
    as_list = FALSE
  )
```

Statistics for the `longitude` variable.
:::

::: {#fig-sm-5-var-distribution-histogram-longitude}
```{r}
#| code-fold: true

data |>
  test_normality(
    col = 'longitude',
    name = 'Longitude (decimal degrees)'
  )
```

Histogram of the `longitude` variable with a kernel density
estimate, along with a quantile-quantile (Q-Q) plot between the
variable and the theoretical quantiles of the normal distribution.
:::

::: {#fig-sm-5-var-distribution-box-plot-longitude}
```{r}
#| code-fold: true

data |>
  plotr:::plot_box_plot(
    col = 'longitude'
  )
```

Boxplot of the `longitude` variable.
:::

### Monthly average global horizontal irradiance (Wh/m²)

::: {#tbl-sm-5-var-distribution-stats-ghi-month}
```{r}
#| code-fold: true

data |>
  stats_summary(
    col = 'ghi_month',
    name = 'Monthly average global horizontal irradiance (Wh/m²)',
    as_list = FALSE
  )
```

Statistics for the `ghi_month` variable.
:::

::: {#fig-sm-5-var-distribution-histogram-ghi-month}
```{r}
#| code-fold: true

data |>
  test_normality(
    col = 'ghi_month',
    name = 'Monthly average global horizontal irradiance (Wh/m²)'
  )
```

Histogram of the `ghi_month` variable with a kernel density
estimate, along with a quantile-quantile (Q-Q) plot between the
variable and the theoretical quantiles of the normal distribution.
:::

::: {#fig-sm-5-var-distribution-box-plot-ghi-month}
```{r}
#| code-fold: true

data |>
  plotr:::plot_box_plot(
    col = 'ghi_month'
  )
```

Boxplot of the `ghi_month` variable.
:::

### Annual average global horizontal irradiance (Wh/m²)

::: {#tbl-sm-5-var-distribution-stats-ghi-annual}
```{r}
#| code-fold: true

data |>
  stats_summary(
    col = 'ghi_annual',
    name = 'Annual average global horizontal irradiance (Wh/m²)',
    as_list = FALSE
  )
```

Statistics for the `ghi_annual` variable.
:::

::: {#fig-sm-5-var-distribution-histogram-ghi-annual}
```{r}
#| code-fold: true

data |>
  test_normality(
    col = 'ghi_annual',
    name = 'Annual average global horizontal irradiance (Wh/m²)'
  )
```

Histogram of the `ghi_annual` variable with a kernel density
estimate, along with a quantile-quantile (Q-Q) plot between the
variable and the theoretical quantiles of the normal distribution.
:::

::: {#fig-sm-5-var-distribution-box-plot-ghi-annual}
```{r}
#| code-fold: true

data |>
  plotr:::plot_box_plot(
    col = 'ghi_annual'
  )
```

Boxplot of the `ghi_annual` variable.
:::

### Daylight on the March equinox (seconds)

::: {#tbl-sm-5-var-distribution-stats-march-equinox-daylight}
```{r}
#| code-fold: true

data |>
  stats_summary(
    col = 'march_equinox_daylight',
    name = 'Daylight on the March equinox (seconds)',
    as_list = FALSE
  )
```

Statistics for the `march_equinox_daylight` variable.
:::

::: {#fig-sm-5-var-distribution-histogram-march-equinox-daylight}
```{r}
#| code-fold: true

data |>
  test_normality(
    col = 'march_equinox_daylight',
    name = 'Daylight on the March equinox (seconds)'
  )
```

Histogram of the `march_equinox_daylight` variable with a kernel density
estimate, along with a quantile-quantile (Q-Q) plot between the
variable and the theoretical quantiles of the normal distribution.
:::

::: {#fig-sm-5-var-distribution-box-plot-march-equinox-daylight}
```{r}
#| code-fold: true

data |>
  plotr:::plot_box_plot(
    col = 'march_equinox_daylight'
  )
```

Boxplot of the `march_equinox_daylight` variable.
:::

### Daylight on the June solstice (seconds)

::: {#tbl-sm-5-var-distribution-stats-june-solstice-daylight}
```{r}
#| code-fold: true

data |>
  stats_summary(
    col = 'june_solstice_daylight',
    name = 'Daylight on the June solstice (seconds)',
    as_list = FALSE
  )
```

Statistics for the `june_solstice_daylight` variable.
:::

::: {#fig-sm-5-var-distribution-histogram-june-solstice-daylight}
```{r}
#| code-fold: true

data |>
  test_normality(
    col = 'june_solstice_daylight',
    name = 'Daylight on the June solstice (seconds)'
  )
```

Histogram of the `june_solstice_daylight` variable with a kernel density
estimate, along with a quantile-quantile (Q-Q) plot between the
variable and the theoretical quantiles of the normal distribution.
:::

::: {#fig-sm-5-var-distribution-box-plot-june-solstice-daylight}
```{r}
#| code-fold: true

data |>
  plotr:::plot_box_plot(
    col = 'june_solstice_daylight'
  )
```

Boxplot of the `june_solstice_daylight` variable.
:::

### Daylight on the December solstice (seconds)

::: {#tbl-sm-5-var-distribution-stats-december-solstice-daylight}
```{r}
#| code-fold: true

data |>
  stats_summary(
    col = 'december_solstice_daylight',
    name = 'Daylight on the December solstice (seconds)',
    as_list = FALSE
  )
```

Statistics for the `december_solstice_daylight` variable.
:::

::: {#fig-sm-5-var-distribution-histogram-december-solstice-daylight}
```{r}
#| code-fold: true

data |>
  test_normality(
    col = 'december_solstice_daylight',
    name = 'Daylight on the December solstice (seconds)'
  )
```

Histogram of the `december_solstice_daylight` variable with a kernel density
estimate, along with a quantile-quantile (Q-Q) plot between the
variable and the theoretical quantiles of the normal distribution.
:::

::: {#fig-sm-5-var-distribution-box-plot-december-solstice-daylight}
```{r}
#| code-fold: true

data |>
  plotr:::plot_box_plot(
    col = 'december_solstice_daylight'
  )
```

Boxplot of the `december_solstice_daylight` variable.
:::
::::
