:::: {.panel-tabset}

```{r}
#| code-fold: true

  plots <- fit_engine |> olsrr::ols_plot_dfbetas(print_plot = FALSE)
  coef_names <- stats::coef(fit_engine) |> names()
```

###### Intercept

::: {#tbl-sm-5-restricted-model-diag-influence-intercept}
```{r}
#| code-fold: true

plots$plots[[1]] +
 ggplot2::labs(title = 'Intercept coefficient')
```

Standardized DFBETAS values for each observation concerning the
`Intercept` coefficient.
:::

###### age

::: {#tbl-sm-5-restricted-model-diag-influence-age}
```{r}
#| code-fold: true

plots$plots[[2]] +
 ggplot2::labs(title = 'age coefficient')
```

Standardized DFBETAS values for each observation concerning the
`age` coefficient.
:::

###### sexMale

::: {#tbl-sm-5-restricted-model-diag-influence-sexmale}
```{r}
#| code-fold: true

plots$plots[[3]] +
 ggplot2::labs(title = 'sexMale coefficient')
```

Standardized DFBETAS values for each observation concerning the
`sexMale` coefficient.
:::

###### longitude

::: {#tbl-sm-5-restricted-model-diag-influence-longitude}
```{r}
#| code-fold: true

plots$plots[[4]] +
 ggplot2::labs(title = 'longitude coefficient')
```

Standardized DFBETAS values for each observation concerning the
`longitude` coefficient.
:::

###### ghi_month

::: {#tbl-sm-5-restricted-model-diag-influence-ghi-month}
```{r}
#| code-fold: true

plots$plots[[5]] +
 ggplot2::labs(title = 'ghi_month coefficient')
```

Standardized DFBETAS values for each observation concerning the
`ghi_month` coefficient.
:::
::::
