:::: {.panel-tabset}
###### MSF~sc~ (Chronotype proxy) (seconds)

::: {#tbl-sm-6-full-model-diag-homoscedasticity-msf-sc}
```{r}
#| code-fold: true

plot <-
  fit |>
  stats::predict(data) |>
  dplyr::mutate(
    .sd_resid =
      fit |>
      parsnip::extract_fit_engine() |>
      stats::rstandard() |>
      abs() |>
      sqrt()
  ) |>
  dplyr::bind_cols(data) |>
  ggplot2::ggplot(ggplot2::aes(msf_sc, .sd_resid)) +
  ggplot2::geom_point() +
  ggplot2::geom_smooth(color = 'red') +
  ggplot2::labs(
    x = 'MSF~sc~ (Chronotype proxy) (seconds)',
    y = latex2exp::TeX('$\\sqrt{|Standardized \\ Residuals|}$')
  )

plot |> print() |> rutils::shush()
```

Relation between `msf_sc` and the model standardized residuals.
:::

###### Age (years)

::: {#tbl-sm-6-full-model-diag-homoscedasticity-age}
```{r}
#| code-fold: true

plot <-
  fit |>
  stats::predict(data) |>
  dplyr::mutate(
    .sd_resid =
      fit |>
      parsnip::extract_fit_engine() |>
      stats::rstandard() |>
      abs() |>
      sqrt()
  ) |>
  dplyr::bind_cols(data) |>
  ggplot2::ggplot(ggplot2::aes(age, .sd_resid)) +
  ggplot2::geom_point() +
  ggplot2::geom_smooth(color = 'red') +
  ggplot2::labs(
    x = 'Age (years)',
    y = latex2exp::TeX('$\\sqrt{|Standardized \\ Residuals|}$')
  )

plot |> print() |> rutils::shush()
```

Relation between `age` and the model standardized residuals.
:::

###### Latitude (decimal degrees)

::: {#tbl-sm-6-full-model-diag-homoscedasticity-longitude}
```{r}
#| code-fold: true

plot <-
  fit |>
  stats::predict(data) |>
  dplyr::mutate(
    .sd_resid =
      fit |>
      parsnip::extract_fit_engine() |>
      stats::rstandard() |>
      abs() |>
      sqrt()
  ) |>
  dplyr::bind_cols(data) |>
  ggplot2::ggplot(ggplot2::aes(longitude, .sd_resid)) +
  ggplot2::geom_point() +
  ggplot2::geom_smooth(color = 'red') +
  ggplot2::labs(
    x = 'Latitude (decimal degrees)',
    y = latex2exp::TeX('$\\sqrt{|Standardized \\ Residuals|}$')
  )

plot |> print() |> rutils::shush()
```

Relation between `longitude` and the model standardized residuals.
:::

###### Longitude (decimal degrees)

::: {#tbl-sm-6-full-model-diag-homoscedasticity-ghi-month}
```{r}
#| code-fold: true

plot <-
  fit |>
  stats::predict(data) |>
  dplyr::mutate(
    .sd_resid =
      fit |>
      parsnip::extract_fit_engine() |>
      stats::rstandard() |>
      abs() |>
      sqrt()
  ) |>
  dplyr::bind_cols(data) |>
  ggplot2::ggplot(ggplot2::aes(ghi_month, .sd_resid)) +
  ggplot2::geom_point() +
  ggplot2::geom_smooth(color = 'red') +
  ggplot2::labs(
    x = 'Longitude (decimal degrees)',
    y = latex2exp::TeX('$\\sqrt{|Standardized \\ Residuals|}$')
  )

plot |> print() |> rutils::shush()
```

Relation between `ghi_month` and the model standardized residuals.
:::

###### Monthly average global horizontal irradiance (Wh/m²)

::: {#tbl-sm-6-full-model-diag-homoscedasticity-latitude}
```{r}
#| code-fold: true

plot <-
  fit |>
  stats::predict(data) |>
  dplyr::mutate(
    .sd_resid =
      fit |>
      parsnip::extract_fit_engine() |>
      stats::rstandard() |>
      abs() |>
      sqrt()
  ) |>
  dplyr::bind_cols(data) |>
  ggplot2::ggplot(ggplot2::aes(latitude, .sd_resid)) +
  ggplot2::geom_point() +
  ggplot2::geom_smooth(color = 'red') +
  ggplot2::labs(
    x = 'Monthly average global horizontal irradiance (Wh/m²)',
    y = latex2exp::TeX('$\\sqrt{|Standardized \\ Residuals|}$')
  )

plot |> print() |> rutils::shush()
```

Relation between `latitude` and the model standardized residuals.
:::
::::
