:::: {.panel-tabset}
### MSF~sc~ (Chronotype proxy) (Seconds)

::: {#tbl-sm-6-var-dist-stats-msf-sc}
```{r}
#| code-fold: true
#| output: asis

data |>
  rutils:::stats_summary(
    col = "msf_sc",
    na_rm = TRUE,
    remove_outliers = FALSE,
    iqr_mult = 1.5,
    hms_format = TRUE,
    threshold = hms::parse_hms("12:00:00"),
    as_list = FALSE
  )
```

[Source: Created by the author.]{.legend}

Statistics for the `msf_sc` variable.
:::

::: {#fig-sm-6-var-dist-charts-hist-qq-plot-msf-sc}
```{r}
#| code-fold: true

data |>
  plotr:::plot_dist(
    col = "msf_sc",
    x_label = "MSF~sc~ (Chronotype proxy) (Seconds)"
  )
```

[Source: Created by the author.]{.legend}

Histogram of the `msf_sc` variable with a kernel density
estimate, along with a quantile-quantile (Q-Q) plot between the
variable and the theoretical quantiles of the normal distribution.
:::

::: {#fig-sm-6-var-dist-charts-box-plot-msf-sc}
```{r}
#| code-fold: true

data |> plotr:::plot_box_plot(col = "msf_sc")
```

[Source: Created by the author.]{.legend}

Box plot of the `msf_sc` variable with jittered data points in black.
:::

### Age (Years)

::: {#tbl-sm-6-var-dist-stats-age}
```{r}
#| code-fold: true
#| output: asis

data |>
  rutils:::stats_summary(
    col = "age",
    na_rm = TRUE,
    remove_outliers = FALSE,
    iqr_mult = 1.5,
    hms_format = TRUE,
    threshold = hms::parse_hms("12:00:00"),
    as_list = FALSE
  )
```

[Source: Created by the author.]{.legend}

Statistics for the `age` variable.
:::

::: {#fig-sm-6-var-dist-charts-hist-qq-plot-age}
```{r}
#| code-fold: true

data |>
  plotr:::plot_dist(
    col = "age",
    x_label = "Age (Years)"
  )
```

[Source: Created by the author.]{.legend}

Histogram of the `age` variable with a kernel density
estimate, along with a quantile-quantile (Q-Q) plot between the
variable and the theoretical quantiles of the normal distribution.
:::

::: {#fig-sm-6-var-dist-charts-box-plot-age}
```{r}
#| code-fold: true

data |> plotr:::plot_box_plot(col = "age")
```

[Source: Created by the author.]{.legend}

Box plot of the `age` variable with jittered data points in black.
:::

### Latitude (Decimal degrees)

::: {#tbl-sm-6-var-dist-stats-latitude}
```{r}
#| code-fold: true
#| output: asis

data |>
  rutils:::stats_summary(
    col = "latitude",
    na_rm = TRUE,
    remove_outliers = FALSE,
    iqr_mult = 1.5,
    hms_format = TRUE,
    threshold = hms::parse_hms("12:00:00"),
    as_list = FALSE
  )
```

[Source: Created by the author.]{.legend}

Statistics for the `latitude` variable.
:::

::: {#fig-sm-6-var-dist-charts-hist-qq-plot-latitude}
```{r}
#| code-fold: true

data |>
  plotr:::plot_dist(
    col = "latitude",
    x_label = "Latitude (Decimal degrees)"
  )
```

[Source: Created by the author.]{.legend}

Histogram of the `latitude` variable with a kernel density
estimate, along with a quantile-quantile (Q-Q) plot between the
variable and the theoretical quantiles of the normal distribution.
:::

::: {#fig-sm-6-var-dist-charts-box-plot-latitude}
```{r}
#| code-fold: true

data |> plotr:::plot_box_plot(col = "latitude")
```

[Source: Created by the author.]{.legend}

Box plot of the `latitude` variable with jittered data points in black.
:::

### Longitude (Decimal degrees)

::: {#tbl-sm-6-var-dist-stats-longitude}
```{r}
#| code-fold: true
#| output: asis

data |>
  rutils:::stats_summary(
    col = "longitude",
    na_rm = TRUE,
    remove_outliers = FALSE,
    iqr_mult = 1.5,
    hms_format = TRUE,
    threshold = hms::parse_hms("12:00:00"),
    as_list = FALSE
  )
```

[Source: Created by the author.]{.legend}

Statistics for the `longitude` variable.
:::

::: {#fig-sm-6-var-dist-charts-hist-qq-plot-longitude}
```{r}
#| code-fold: true

data |>
  plotr:::plot_dist(
    col = "longitude",
    x_label = "Longitude (Decimal degrees)"
  )
```

[Source: Created by the author.]{.legend}

Histogram of the `longitude` variable with a kernel density
estimate, along with a quantile-quantile (Q-Q) plot between the
variable and the theoretical quantiles of the normal distribution.
:::

::: {#fig-sm-6-var-dist-charts-box-plot-longitude}
```{r}
#| code-fold: true

data |> plotr:::plot_box_plot(col = "longitude")
```

[Source: Created by the author.]{.legend}

Box plot of the `longitude` variable with jittered data points in black.
:::

### Monthly average global horizontal irradiance (Wh/m²)

::: {#tbl-sm-6-var-dist-stats-ghi-month}
```{r}
#| code-fold: true
#| output: asis

data |>
  rutils:::stats_summary(
    col = "ghi_month",
    na_rm = TRUE,
    remove_outliers = FALSE,
    iqr_mult = 1.5,
    hms_format = TRUE,
    threshold = hms::parse_hms("12:00:00"),
    as_list = FALSE
  )
```

[Source: Created by the author.]{.legend}

Statistics for the `ghi_month` variable.
:::

::: {#fig-sm-6-var-dist-charts-hist-qq-plot-ghi-month}
```{r}
#| code-fold: true

data |>
  plotr:::plot_dist(
    col = "ghi_month",
    x_label = "Monthly average global horizontal irradiance (Wh/m²)"
  )
```

[Source: Created by the author.]{.legend}

Histogram of the `ghi_month` variable with a kernel density
estimate, along with a quantile-quantile (Q-Q) plot between the
variable and the theoretical quantiles of the normal distribution.
:::

::: {#fig-sm-6-var-dist-charts-box-plot-ghi-month}
```{r}
#| code-fold: true

data |> plotr:::plot_box_plot(col = "ghi_month")
```

[Source: Created by the author.]{.legend}

Box plot of the `ghi_month` variable with jittered data points in black.
:::
::::
