:::: {.panel-tabset}
###### MSF~sc~ (Chronotype proxy) (seconds)

::: {#tbl-sm-5-fm-var-homoscedasticity-msf-sc}
```{r}
#| code-fold: true

fit |>
  plotr:::plot_homoscedasticity(
    data = data,
    col = "msf_sc",
    x_label = "MSF~sc~ (Chronotype proxy) (seconds)"
  )
```

Relation between `msf_sc` and the model standardized residuals.
:::

###### Age (years)

::: {#tbl-sm-5-fm-var-homoscedasticity-age}
```{r}
#| code-fold: true

fit |>
  plotr:::plot_homoscedasticity(
    data = data,
    col = "age",
    x_label = "Age (years)"
  )
```

Relation between `age` and the model standardized residuals.
:::

###### Longitude (decimal degrees)

::: {#tbl-sm-5-fm-var-homoscedasticity-longitude}
```{r}
#| code-fold: true

fit |>
  plotr:::plot_homoscedasticity(
    data = data,
    col = "longitude",
    x_label = "Longitude (decimal degrees)"
  )
```

Relation between `longitude` and the model standardized residuals.
:::

###### Monthly average global horizontal irradiance (Wh/m²)

::: {#tbl-sm-5-fm-var-homoscedasticity-ghi-month}
```{r}
#| code-fold: true

fit |>
  plotr:::plot_homoscedasticity(
    data = data,
    col = "ghi_month",
    x_label = "Monthly average global horizontal irradiance (Wh/m²)"
  )
```

Relation between `ghi_month` and the model standardized residuals.
:::

###### Annual average global horizontal irradiance (Wh/m²)

::: {#tbl-sm-5-fm-var-homoscedasticity-ghi-annual}
```{r}
#| code-fold: true

fit |>
  plotr:::plot_homoscedasticity(
    data = data,
    col = "ghi_annual",
    x_label = "Annual average global horizontal irradiance (Wh/m²)"
  )
```

Relation between `ghi_annual` and the model standardized residuals.
:::

###### Daylight on the March equinox (seconds)

::: {#tbl-sm-5-fm-var-homoscedasticity-march-equinox-daylight}
```{r}
#| code-fold: true

fit |>
  plotr:::plot_homoscedasticity(
    data = data,
    col = "march_equinox_daylight",
    x_label = "Daylight on the March equinox (seconds)"
  )
```

Relation between `march_equinox_daylight` and the model standardized residuals.
:::

###### Daylight on the June solstice (seconds)

::: {#tbl-sm-5-fm-var-homoscedasticity-june-solstice-daylight}
```{r}
#| code-fold: true

fit |>
  plotr:::plot_homoscedasticity(
    data = data,
    col = "june_solstice_daylight",
    x_label = "Daylight on the June solstice (seconds)"
  )
```

Relation between `june_solstice_daylight` and the model standardized residuals.
:::

###### Daylight on the December solstice (seconds)

::: {#tbl-sm-5-fm-var-homoscedasticity-december-solstice-daylight}
```{r}
#| code-fold: true

fit |>
  plotr:::plot_homoscedasticity(
    data = data,
    col = "december_solstice_daylight",
    x_label = "Daylight on the December solstice (seconds)"
  )
```

Relation between `december_solstice_daylight` and the model standardized residuals.
:::
::::
