:::: {.panel-tabset}
###### MSF~sc~ (Chronotype proxy) (seconds)

::: {#tbl-sm-6-fm-var-homoscedasticity-msf-sc}
```{r}
#| code-fold: true

fit |>
  plotr:::plot_homoscedasticity(
    data = data,
    col = "msf_sc",
    x_label = "MSF~sc~ (Chronotype proxy) (seconds)"
  )
```

Relation between `msf_sc` and the model standardized residuals.
:::

###### Age (years)

::: {#tbl-sm-6-fm-var-homoscedasticity-age}
```{r}
#| code-fold: true

fit |>
  plotr:::plot_homoscedasticity(
    data = data,
    col = "age",
    x_label = "Age (years)"
  )
```

Relation between `age` and the model standardized residuals.
:::

###### Latitude (decimal degrees)

::: {#tbl-sm-6-fm-var-homoscedasticity-longitude}
```{r}
#| code-fold: true

fit |>
  plotr:::plot_homoscedasticity(
    data = data,
    col = "longitude",
    x_label = "Latitude (decimal degrees)"
  )
```

Relation between `longitude` and the model standardized residuals.
:::

###### Longitude (decimal degrees)

::: {#tbl-sm-6-fm-var-homoscedasticity-ghi-month}
```{r}
#| code-fold: true

fit |>
  plotr:::plot_homoscedasticity(
    data = data,
    col = "ghi_month",
    x_label = "Longitude (decimal degrees)"
  )
```

Relation between `ghi_month` and the model standardized residuals.
:::

###### Monthly average global horizontal irradiance (Wh/m²)

::: {#tbl-sm-6-fm-var-homoscedasticity-latitude}
```{r}
#| code-fold: true

fit |>
  plotr:::plot_homoscedasticity(
    data = data,
    col = "latitude",
    x_label = "Monthly average global horizontal irradiance (Wh/m²)"
  )
```

Relation between `latitude` and the model standardized residuals.
:::
::::
