:::: {.panel-tabset}

```{r}
#| code-fold: true

  plots <- fit_engine |> olsrr::ols_plot_dfbetas(print_plot = FALSE)
  coef_names <- stats::coef(fit_engine) |> names()
```

###### Intercept

::: {#tbl-sm-5-fm-coef-dfbetas-intercept}
```{r}
#| code-fold: true

plots$plots[[1]] +
ggplot2::labs(title = "Intercept coefficient")
```

Standardized DFBETAS values for each observation concerning the
`Intercept` coefficient.
:::

###### age

::: {#tbl-sm-5-fm-coef-dfbetas-age}
```{r}
#| code-fold: true

plots$plots[[2]] +
ggplot2::labs(title = "age coefficient")
```

Standardized DFBETAS values for each observation concerning the
`age` coefficient.
:::

###### sexMale

::: {#tbl-sm-5-fm-coef-dfbetas-sexmale}
```{r}
#| code-fold: true

plots$plots[[3]] +
ggplot2::labs(title = "sexMale coefficient")
```

Standardized DFBETAS values for each observation concerning the
`sexMale` coefficient.
:::

###### longitude

::: {#tbl-sm-5-fm-coef-dfbetas-longitude}
```{r}
#| code-fold: true

plots$plots[[4]] +
ggplot2::labs(title = "longitude coefficient")
```

Standardized DFBETAS values for each observation concerning the
`longitude` coefficient.
:::

###### ghi_month

::: {#tbl-sm-5-fm-coef-dfbetas-ghi-month}
```{r}
#| code-fold: true

plots$plots[[5]] +
ggplot2::labs(title = "ghi_month coefficient")
```

Standardized DFBETAS values for each observation concerning the
`ghi_month` coefficient.
:::

###### ghi_annual

::: {#tbl-sm-5-fm-coef-dfbetas-ghi-annual}
```{r}
#| code-fold: true

plots$plots[[6]] +
ggplot2::labs(title = "ghi_annual coefficient")
```

Standardized DFBETAS values for each observation concerning the
`ghi_annual` coefficient.
:::

###### march_equinox_daylight

::: {#tbl-sm-5-fm-coef-dfbetas-march-equinox-daylight}
```{r}
#| code-fold: true

plots$plots[[7]] +
ggplot2::labs(title = "march_equinox_daylight coefficient")
```

Standardized DFBETAS values for each observation concerning the
`march_equinox_daylight` coefficient.
:::

###### june_solstice_daylight

::: {#tbl-sm-5-fm-coef-dfbetas-june-solstice-daylight}
```{r}
#| code-fold: true

plots$plots[[8]] +
ggplot2::labs(title = "june_solstice_daylight coefficient")
```

Standardized DFBETAS values for each observation concerning the
`june_solstice_daylight` coefficient.
:::

###### december_solstice_daylight

::: {#tbl-sm-5-fm-coef-dfbetas-december-solstice-daylight}
```{r}
#| code-fold: true

plots$plots[[9]] +
ggplot2::labs(title = "december_solstice_daylight coefficient")
```

Standardized DFBETAS values for each observation concerning the
`december_solstice_daylight` coefficient.
:::
::::
