:::: {.panel-tabset}

```{r}
#| code-fold: true

  plots <- fit_engine |> olsrr::ols_plot_dfbetas(print_plot = FALSE)
  coef_names <- stats::coef(fit_engine) |> names()
```

###### Intercept

::: {#tbl-sm-6-fm-coef-dfbetas-intercept}
```{r}
#| code-fold: true

plots$plots[[1]] +
ggplot2::labs(title = "Intercept coefficient")
```

Standardized DFBETAS values for each observation concerning the
`Intercept` coefficient.
:::

###### age

::: {#tbl-sm-6-fm-coef-dfbetas-age}
```{r}
#| code-fold: true

plots$plots[[2]] +
ggplot2::labs(title = "age coefficient")
```

Standardized DFBETAS values for each observation concerning the
`age` coefficient.
:::

###### sexMale

::: {#tbl-sm-6-fm-coef-dfbetas-sexmale}
```{r}
#| code-fold: true

plots$plots[[3]] +
ggplot2::labs(title = "sexMale coefficient")
```

Standardized DFBETAS values for each observation concerning the
`sexMale` coefficient.
:::

###### longitude

::: {#tbl-sm-6-fm-coef-dfbetas-longitude}
```{r}
#| code-fold: true

plots$plots[[4]] +
ggplot2::labs(title = "longitude coefficient")
```

Standardized DFBETAS values for each observation concerning the
`longitude` coefficient.
:::

###### ghi_month

::: {#tbl-sm-6-fm-coef-dfbetas-ghi-month}
```{r}
#| code-fold: true

plots$plots[[5]] +
ggplot2::labs(title = "ghi_month coefficient")
```

Standardized DFBETAS values for each observation concerning the
`ghi_month` coefficient.
:::

###### latitude

::: {#tbl-sm-6-fm-coef-dfbetas-latitude}
```{r}
#| code-fold: true

plots$plots[[6]] +
ggplot2::labs(title = "latitude coefficient")
```

Standardized DFBETAS values for each observation concerning the
`latitude` coefficient.
:::
::::
