## Chapter 5

<!-- $\text{`{r} weighted_data |> dplyr::pull(msf_sc) |> mean() |> hms::as_hms() |> lubritime::round_time()`}$ -->

<!-- $\text{`{r} weighted_data |> dplyr::pull(msf_sc) |> stats::sd() |> hms::as_hms() |> lubritime::round_time()`}$ -->

The local time of the sleep-corrected midpoint between sleep onset and sleep end on work-free days (MSFsc), which serves as the MCTQ proxy for measuring chronotype, had an overall mean of  $\text{04:28:41}$ and a standard deviation of $\text{01:26:13}$. The distribution curve is shown in @fig-chapter-5-chronotype-distribution.

<!-- $\text{`{r} weighted_data |> dplyr::pull(msf_sc) |> mean() %>% `+`(hms::parse_hm("04:00")) |> hms::as_hms() |> lubritime::round_time()`}$ -->

This represents the midsleep point for Brazilian subjects with an intermediate or average chronotype. Considering the 7–9 hours of sleep recommended for healthy adults by the American Academy of Sleep Medicine (AASM) [@watson2015], one could imagine that this average individual, in the absence of social restrains, would typically wake up at approximately $\text{08:28:41}$.

## Methods - Sample

```{r}
#| include: false

data <- 
  anonymized_data |> 
  dplyr::filter(country == "Brazil")
```

```{r}
#| include: false

sex_per <- 
  weighted_data |> 
  dplyr::summarise(
    n = dplyr::n(), 
    .by = sex
  ) |>
  dplyr::mutate(n_per = (n / sum(n)) * 100)

male_per <- 
  sex_per |> 
  dplyr::filter(sex == "Male") |>
  dplyr::pull(n_per)

female_per <- 
  sex_per |> 
  dplyr::filter(sex == "Female") |>
  dplyr::pull(n_per)
```

Based on 2017 data from the Brazilian Institute of Geography and Statistics's ([IBGE](https://www.ibge.gov.br/)) Continuous National Household Sample Survey ([PNAD Contínua](https://www.ibge.gov.br/estatisticas/sociais/trabalho/17270-pnad-continua.html)) [@ibgee], Brazil had $51.919\%$ of females and $48.081\%$ of males with an age equal to or greater than 18 years old. The full sample is skewed for female subjects, with `{r} format_to_md_latex(female_per, "\\%")` of females and `{r} format_to_md_latex(male_per, "\\%")` of male subjects.

```{r}
#| include: false

total_age_mean <- 
  data |>
  dplyr::pull(age) |>
  mean(na.rm = TRUE)

total_age_sd <- 
  data |>
  dplyr::pull(age) |>
  stats::sd(na.rm = TRUE)

age_mean_sd <- 
  data |> 
  dplyr::summarise(
    mean = mean(age, na.rm = TRUE),
    sd = stats::sd(age, na.rm = TRUE),
    .by = c("sex")
  )

male_age_mean <- 
  age_mean_sd |> 
  dplyr::filter(sex == "Male") |>
  dplyr::pull(mean)

male_age_sd <- 
  age_mean_sd |> 
  dplyr::filter(sex == "Male") |>
  dplyr::pull(sd)

female_age_mean <- 
  age_mean_sd |> 
  dplyr::filter(sex == "Female") |>
  dplyr::pull(mean)

female_age_sd <- 
  age_mean_sd |> 
  dplyr::filter(sex == "Female") |>
  dplyr::pull(sd)
```

Subjects' mean age is `{r} format_to_md_latex(total_age_mean)` years ($\text{SD} = 9.252$; $\text{Max.} = 58.786$). Female subjects have a mean age of $31.787$ years ($\text{SD} = 9.364$; $\text{Max.} = 58.786$) and male subjects $32.464$ years ($\text{SD} = 9.012$; $\text{Max.} = 58.772$). For comparison, based on the 2017 IBGE-PNAD data [@ibgee], Brazil’s population with an age equal to or greater than $18$ years old had a mean age of $44.277$ years ($\text{SD} = 17.221$), with a mean age of $44.987$ years ($\text{SD} = 17.511$) for female subjects and a mean age of $43.499$ years ($\text{SD} = 16.864$) for male subjects.

Considering the five regions of Brazil, the sample is mostly skewed for the Southeast, the most populated region. According to Brazil’s 2022 census [@ibge2022], the Southeast region is home to $41.784\%$ of Brazil’s population, followed by the Northeast ($26.910\%$), South ($14.741\%$), North ($8.544\%$), and Midwest ($8.021\%$) regions. $62.454\%$ of the sample is located in the Southeast region, $11.797\%$ in the Northeast, $17.861\%$ in the South, $1.682\%$ in the North, and $6.205\%$ in the Midwest region. Note that a lack of subjects in the North and Midwest region is justified by the sample timezone inclusion criteria (UTC-3).

The sample latitudinal range was $30.211$ decimal degrees ($\text{Min.} = -30.109$; $\text{Max.} = 0.10177$) with a longitudinal span of $16.378$ decimal degrees ($\text{Min.} = -51.342$; $\text{Max.} = -34.964$). For comparison, Brazil has a latitudinal range of $39.024$ decimal degrees ($\text{Min.} = -33.752$; $\text{Max.} = 5.2719$) and a longitudinal span of $39.198$ decimal degrees ($\text{Min.} = -34.793$; $\text{Max.} = -73.991$). 
