# Abbreviations {.unnumbered}

```{r}
#| label: setup
#| echo: false
#| output: false

library(here)
source(here::here("R/quarto-setup.R"))
```

```{r}
#| echo: false
#| output: asis

# library(rutils)

rutils:::quarto_status("polishing")
```

* __~F~__: Subscript indicating a relation with work-free days.
* __~W~__: Subscript indicating a relation with workdays.
* __BT__: Local time of going to bed.
* __FD__: Number of work-free days per week.
* __GU__: Local time of getting out of bed.
* __HO__: Horne & Ostberg's morningness-eveningness questionnaire (same as _MEQ_).
* __LE__: Light exposure.
* __LE~week~__: Average weekly light exposure.
* __MCTQ__: Munich ChronoType Questionnaire.
* __MCTQ^PT^__: Portuguese version of the MCTQ.
* __MEQ__: Morningness-Eveningness Questionnaire.
* __MSF__: Local time of mid-sleep on work-free days.
* __MSF~sc~__: Chronotype proxy. The midpoint between sleep onset and sleep end on work-free days. A sleep correction (~SC~) is made when a possible sleep compensation related to a lack of sleep on workdays is identified.
* __MSW__: Local time of mid-sleep on workdays.
* __PRC__: Phase response curve.
* __SD__: Sleep duration.
* __SD~week~__: Average weekly sleep duration.
* __SE__: Local time of sleep end.
* __SI__: "Sleep inertia". Despite the name, this abbreviation represents the time the respondent takes to get up after sleep end. It is used this way by the MCTQ authors.
* __SJL__: Absolute social jetlag.
* __SJL~rel~__: Relative social jetlag.
* __SJL~sc~__: Jankowski's sleep-corrected social jetlag.
* __SJL~sc-rel~__: Jankowski's relative sleep-corrected social jetlag.
* __Sloss~week~__: Weekly sleep loss.
* __SO__: Local time of sleep onset.
* __Slat__: Sleep latency or time to fall asleep after preparing to sleep on workdays.
* __SPrep__: Local time of preparing to sleep.
* __TBT__: Total time in bed.
* __WD__: Number of workdays per week.
