<!-- %:::% .common h1 begin %:::% -->
# Ecology of sleep and circadian phenotypes of the brazilian population
<!-- %:::% .common h1 end %:::% -->

```{r}
#| label: setup
#| echo: false
#| output: false

library(here)
source(here::here("R/quarto-setup.R"))
```

```{r}
#| echo: false
#| output: asis

# library(rutils)

rutils:::quarto_status("drafting")
```

::: {.callout-note}
## Target

1. [Chronobiology International](https://www.tandfonline.com/action/authorSubmission?show=instructions&journalCode=icbi20) ([IF 2022: 2.8/JCR](https://jcr.clarivate.com/jcr/) | [A1/2017-2020](https://sucupira.capes.gov.br/sucupira/public/consultas/coleta/veiculoPublicacaoQualis/listaConsultaGeralPeriodicos.jsf)).
2. [Journal of Biological Rhythms](https://journals.sagepub.com/author-instructions/JBR) ([IF 2022: 3.5/JCR](https://jcr.clarivate.com/jcr/) | [A2/2017-2020](https://sucupira.capes.gov.br/sucupira/public/consultas/coleta/veiculoPublicacaoQualis/listaConsultaGeralPeriodicos.jsf)).
:::

::: {.callout-note}
The following study was conducted by Daniel Vartanian (DV), Mario Pedrazzoli (MP), and Camilo Rodrigues Neto (CR).

::: {.content-visible when-format="pdf"}
\vspace{5pt}
:::

__DV__ conceived the study, contributed with the design, implementation, statistical analysis and authored the manuscript. __CR__ contributed as a science adviser and reviewer. __DV__ and __MP__ were responsible for data collection.  All authors actively participated in discussions regarding the results and contributed to the final manuscript.

::: {.content-visible when-format="pdf"}
\vspace{5pt}
:::

_Future reference_: Vartanian, D., Pedrazzoli, M., & Rodrigues Neto, C. (2024). Ecology of sleep and circadian phenotypes of the Brazilian population. _Chronobiology International_.
:::
