<!-- %:::% .common h1 begin %:::% -->
# The {mctq} R package
<!-- %:::% .common h1 end %:::% -->

```{r}
#| label: setup
#| include: false

source(here::here("R/_setup.R"))
```

```{r}
#| echo: false
#| output: asis

rutils:::quarto_status(
  type = "drafting",
  of_what = "of this thesis",
  latex_parskip = "\\microskip"
  )
```

::: {.callout-note}
## Target journal

1. [Journal of Statistical Software](https://www.jstatsoft.org/authors) ([IF 2022: 5.8/JCR](https://jcr.clarivate.com/jcr/) | [A1/2017-2020](https://sucupira.capes.gov.br/sucupira/public/consultas/coleta/veiculoPublicacaoQualis/listaConsultaGeralPeriodicos.jsf)).
2. [Journal of Open Source Software](https://joss.readthedocs.io/en/latest/submitting.html) ([B1/2017-2020](https://sucupira.capes.gov.br/sucupira/public/consultas/coleta/veiculoPublicacaoQualis/listaConsultaGeralPeriodicos.jsf)).
:::

::: {.callout-note}
The following study was conducted by Daniel Vartanian (__DV__), Ana Amélia Benedito-Silva (__AA__), Mario Pedrazzoli (__MP__), and Camilo Rodrigues Neto (__CR__).

\microskip

__DV__ contributed to the conception, design, coding, and implementation of the software. __AA__, __MP__, and __CR__ served as scientific advisors and reviewers. __DV__ authored the manuscript. All authors discussed the results and revised the final manuscript.

\microskip

_Future reference_: Vartanian, D., Benedito-Silva, A. A., Pedrazzoli, M., & Rodrigues Neto, C. (2024). {mctq}: tools to process the Munich ChronoType Questionnaire (MCTQ). _Journal of Statistical Software_.
:::
