<!-- %:::% .common h1 begin %:::% -->
# Conclusion {#sec-conclusion}
<!-- %:::% .common h1 end %:::% -->

```{r}
#| label: setup
#| include: false

source(here::here("R", "_setup.R"))
```

```{r}
#| include: false

source(here::here("R", "utils-md.R"))
```

According to Popper, the aim of science is to provide "satisfactory explanations of whatever strikes us as being in need of explanation" [@popper1979, p. 193]. This study, using what is arguably the largest dataset on chronotype collected within a single time zone, balanced to reflect population proportions at the time of data collection, found no support for the latitude hypothesis. This result contributes meaningful evidence to the understanding of circadian rhythm regulation, offering a clear and satisfactory answer to the central research question of this thesis regarding human populations: _Is latitude associated with chronotype?_ The answer, at present, is **[No]{.brandorange}**.

The development of this work was driven by a commitment to objectivity and scientific rigor, a path that often diverges from the expectations commonly placed on young researchers. The temptation to align with mainstream ideas for easier acceptance within the scientific community — or to increase the likelihood of publication — is ever-present. However, it is crucial to emphasize that this research does not stem from unreasonable skepticism. Instead, it reflects the perspective of a young scientist dedicated to the scientific endeavor and its foundational principles.

## Limitations

While this study provides valuable insights, it is essential to acknowledge certain limitations that may influence the interpretation of the findings.

First, the data collection occurred predominantly during a single week in spring, as summer approached, which limited the photoperiod variability between regions. A better approach would involve data collection across different seasons, particularly during winter, when photoperiod differences are more pronounced between equatorial and polar regions.

Additionally, the use of the Munich Chronotype Questionnaire (MCTQ), while a validated instrument, introduces the potential for recall and social desirability biases inherent to self-reported measures. However, the large sample size likely mitigates these biases, as predicted by the law of large numbers [@degroot2012, p. 352]. Furthermore, at the time of data collection, the MCTQ had not yet been officially validated in Portuguese (this was only introduced in 2020 by @reis2020a), which may have introduced minor inconsistencies, though its nature as a sleep log suggests this impact was minimal.

Another factor to consider is the timing of data collection relative to the start of Daylight Saving Time (DST) in Brazil. On the day data collection commenced (October 15th, 2017 – `{r} res_vars$pr_analysis_sample_per_nrow_2017_10_15 |> format_to_md_latex(after = "\\%")` of the data used in this analysis were collected on this day), a significant portion of respondents adjusted their clocks forward by one hour. While this could theoretically influence their responses, the questions were specifically designed to capture daily routines, which were not affected by the DST adjustment at that moment. Furthermore, any potential effect of DST would likely strengthen the latitude hypothesis; however, this was not supported by the data.

These limitations, while noteworthy, do not undermine the study's findings but rather highlight areas for refinement in future research.

## Directions for Future Research

This thesis proposed using a global modeling approach to investigate the latitude-chronotype relationship. However, as demonstrated by the results of this study and others, no significant effect of latitude on chronotype was identified. That said, it remains possible that if such a phenomenon exists, it could be captured through a localized approach, such as agent-based modeling. This approach would simulate an environment where agents are exposed to varying light levels, while accounting for their endogenous rhythms and the circadian clock's phase-response curve to light. The data from this thesis could serve to calibrate and validate this model.

<!-- textual end -->

```{=latex}
\postextual

\begingroup
\renewcommand{\baselinestretch}{1}
\setcounter{footnote}{0}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\printbibliography[heading=bibheading]
\endgroup

\tocskipone
\tocprintchapternonum
\addcontentsline{toc}{chapter}{\newbibname}
```
