<!-- %:::% .common h1 begin %:::% -->
# Rule-based model of the 24h light/dark entrainment phenomenon
<!-- %:::% .common h1 end %:::% -->

```{r}
#| label: setup
#| echo: false
#| output: false

library(here)
source(here::here("R/quarto-setup.R"))
```

```{r}
#| echo: false
#| output: asis

# library(rutils)

rutils:::quarto_status("drafting")
```

::: {.callout-note}
## Target

1. [Journal of Open Source Software](https://joss.readthedocs.io/en/latest/submitting.html) ([B1/2017-2020](https://sucupira.capes.gov.br/sucupira/public/consultas/coleta/veiculoPublicacaoQualis/listaConsultaGeralPeriodicos.jsf)).
:::

::: {.callout-note}
The following study was conducted by Daniel Vartanian (DV) and Camilo Rodrigues Neto (CR).

::: {.content-visible when-format="pdf"}
\vspace{5pt}
:::

__DV__ was responsible for the design and software implementation. __CR__ contributed as a science adviser and reviewer. __DV__ wrote the manuscript. All authors discussed the results and revised the final manuscript.

::: {.content-visible when-format="pdf"}
\vspace{5pt}
:::

_Future reference_: Vartanian, D, & Rodrigues Neto, C. (2024). {entrainment}: A rule-based model of the 24h light/dark cycle entrainment phenomenon. _Journal of Open Source_.
:::
