<!-- %:::% .common h1 begin %:::% -->
# Conclusion {#sec-conclusion}
<!-- %:::% .common h1 end %:::% -->

```{r}
#| label: setup
#| include: false

source(here::here("R", "_setup.R"))
```

```{r}
#| include: false

source(here::here("R", "utils-md.R"))
```

::: {.content-visible when-format="html"}
> Every genuine test of a theory is an attempt to falsify it, or to refute it [@popper2002].

> I suggest that it is the aim of science to find satisfactory explanations, of whatever strikes us as being in need of explanation [@popper1979a, p. 193].
:::

::: {.content-visible when-format="pdf"}
```{=latex}
\epigraph{Every genuine test of a theory is an attempt to falsify it, or to refute it.}{--- \textcite{popper2002}}
\vspace*{-1\baselineskip}
\epigraph{I suggest that it is the aim of science to find satisfactory explanations, of whatever strikes us as being in need of explanation.}{--- \textcite[193]{popper1979a}}
% \vspace*{1\baselineskip}
```
:::

The preceding chapters have presented a comprehensive examination of evidence and analyses pertaining to the latitude hypothesis in chronobiology. At this juncture, it is essential to return to the fundamental research question that motivated this thesis: *Is latitude associated with chronotype?*

This study, using what is arguably one of the largest datasets on chronotype, found no evidence supporting the latitude hypothesis in humans.

Current evidence does not support the latitude hypothesis, and some claims made in its favor warrant further examination (see Supplemental Materials). Nevertheless, the hypothesis is often cited in chronobiology research as if it were well-established. While this study does not definitively refute it, the relationship between latitude and chronotype should remain an open scientific question until strong evidence substantiates it.

## Strengths

Testing any theory or hypothesis is inherently challenging and often sparks debate. In this context, it is important to acknowledge the significant strengths of this study:

- **Large, focused sample**: One of the largest chronotype datasets ever collected within a single time zone (`{r} res_vars$ea_analysis_sample_n |> groomr:::format_to_md_latex(before = "n = ")`).
- **Minimal photoperiod variability**: Data were collected over a single spring week as summer approached (October 15–21, 2017), effectively controlling for seasonal variations in daylight exposure across regions.
- **Broad latitudinal range**: The sample spans a considerable range (`{r} res_vars$ea_analysis_sample_lat_lon$latitude[3] |> groomr:::format_to_md_latex(after = "°", max_digits = 3)`).
- **Population balancing**: The dataset was adjusted to reflect population proportions at the time of collection.
- **Rigorous statistical analysis**: The study accounted for confounders and incorporated a practical significance threshold.
- **Full reproducibility**: All analyses, from raw data processing through effect size estimation, follow transparent and reproducible protocols available for verification.

It is also worth noting that, despite methodological improvements, this study is grounded in the same core principles and variables that underpin the latitude hypothesis. Consequently, any critique of its methods should be considered in light of the methods used by studies that support the hypothesis.

## Limitations

While this study provides valuable evidence, certain limitations should be acknowledged, as they may influence the interpretation of the findings:

- **Self-report biases**: The Munich Chronotype Questionnaire (MCTQ), despite being a validated instrument, is subject to recall and social desirability biases inherent in self-reported measures. The large sample size likely mitigates these biases, as suggested by the law of large numbers [@degroot2012a, p. 352].
- **Questionnaire validation in Portuguese**: At the time of data collection, the MCTQ had not been officially validated in Portuguese (a process completed in 2020 by @reis2020a), potentially introducing minor inconsistencies. However, given its sleep log format, the impact is expected to be minimal.
- **Daylight Saving Time (DST)**: The timing of data collection coincided with the onset of Daylight Saving Time (DST) in Brazil. On October 15th, 2017, the day data collection commenced (`{r} res_vars$pr_analysis_sample_per_nrow_2017_10_15 |> groomr:::format_to_md_latex(after = "\\%")` of the data used in this analysis were collected on this day), a significant portion of respondents adjusted their clocks forward by one hour. Although this adjustment could theoretically influence responses, the questions were designed to capture daily routines that were not directly affected by the DST shift. Furthermore, if DST had any effect, it would have been expected to bolster the latitude hypothesis; yet, this was not supported by the data.
- **Latitudinal range**: The latitudinal range of `{r} res_vars$ea_analysis_sample_lat_lon$latitude[3] |> groomr:::format_to_md_latex(after = "°", max_digits = 3)`, while substantial, could be questioned as potentially insufficient to detect latitude effects on chronotype. However, the absence of a meaningful association within this range suggests that any such effect, if present, would be minimal.
- **Unexamined confounders**: Although key confounders were controlled for, some variables remained unexamined (e.g., socioeconomic status, urbanization, social timing). While including these variables might enhance model precision, it could also introduce multicollinearity and overfitting concerns. In the context of the tested hypothesis, the exclusion of these additional predictors is unlikely to alter the overall conclusions.
- **Chronotype measurement**: The study used sleep as a proxy for measuring chronotype, leveraging its underlying circadian process. However, sleep is not the only marker of circadian phenotypes. More precise methods, such as Dim Light Melatonin Onset (DLMO) [@ruiz2020], can provide a direct measure of circadian phase. Yet, these approaches are often more invasive and costly, limiting both sample size and the generalizability of findings.
- **Generalizability**: Finally, the study's generalizability is limited to the Brazilian population. However, since the latitude hypothesis's underlying principles are not geographically constrained, these findings provide valuable insights for future research in other regions.

These limitations, while important to consider, do not undermine the study’s findings; rather, they highlight areas where future research might further refine our understanding.

## Directions for Future Research

This thesis proposed using a global modeling approach to investigate the latitude-chronotype relationship. As demonstrated by the results of this study and others, no meaningful effect of latitude on chronotype was identified. That said, it remains possible that if such a phenomenon exists, it could be captured through a localized approach, such as agent-based modeling. This approach would simulate an environment where agents are exposed to varying light levels, while accounting for their endogenous rhythms and the circadian clock's phase-response curve to light. The data from this thesis could serve to calibrate and validate this model.

<!-- textual end -->

```{=latex}
\postextual

\begingroup
\renewcommand{\baselinestretch}{1}
\setcounter{footnote}{0}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\printbibliography[heading=bibheading]
\endgroup

\tocskipone
\tocprintchapternonum
\addcontentsline{toc}{chapter}{\newbibname}
```
