<!-- %:::% index begin %:::% -->
# Introduction

```{r}
#| label: setup
#| include: false

source(here::here("R/_setup.R"))
```

```{r}
#| label: status
#| echo: false
#| output: asis

rutils:::quarto_status(
  type = "polishing",
  of_what = "of this thesis",
  latex_parskip = "\\microskip"
  )
```

<!-- %:::% index pdf begin %:::% -->
There has been a long-standing debate in the chronobiology community regarding the relationship between latitude and circadian phenotypes (chronotypes) [@pittendrigh1991; @zerbini2021; @skeldon2021], with many believing this association is already well-established. This hypothesis stems from the varying amounts of solar radiation experienced by populations across different latitudes. Since light exposure is a primary zeitgeber — a periodic environmental cue that influences or regulates biological rhythms — such variations, along with temperature variations, are thought to create noticeable differences in chronotype distributions globally. This thesis investigates the so-called latitude or environmental hypothesis in human circadian phenotypes, addressing the question: *Is latitude associated with chronotype?*

The central hypothesis is that *latitude is associated with human chronotype distributions*, with populations closer to the equator exhibiting, on average, a shorter or more morning-oriented circadian phenotype compared to those residing near the poles [@hut2013; @leocadio-miguel2014; @leocadio-miguel2017; @pittendrigh1991; @randler2008; @randler2017; @roenneberg2003]. The primary objective is to model and test this hypothesis by critically examining whether a significant association and effect size exist between latitude and circadian phenotypes in the Brazilian population.

This project came to be by a healthy debate that I had with my ex-supervisor about this hypothesis, primarily because of the results he others chronobiology researches published on the matter in 2017 on Nature's Scientific Report journal [@leocadio-miguel2017]. In this paper, the author conclude, as the theory seemes to suggest, that there is a significant association between latitude and chronotype in the Brazilian population. However, the results were not as clear as the author suggested, and the methodology used to test the hypothesis was not the most appropriate. This thesis aims to revisit this hypothesis using a more robust statistical approach and a larger dataset to provide a more accurate and reliable answer to the research question.

In the following chapters, the latitude hypothesis is explored using Popper's hypothetical-deductive method [@popper1979] and an improved approach to Null Hypothesis Significance Testing (NHST), rooted in the original Neyman-Pearson framework for data testing [@neyman1928; @neyman1928a; @perezgonzalez2015]. This exploration involves a series of analyses conducted on a large dataset comprising $120,265$ individuals, collected from the Brazilian population in 2017. The dataset is based on the Munich Chronotype Questionnaire (MCTQ) [@roenneberg2003; @roenneberg2012a], and contains information on sleep habits and demographic characteristics from various regions of Brazil.

The analyses utilize nested multiple regression models to evaluate the additional variance explained and the effect size when latitude is included as a predictor of chronotype. The results are then compared with those obtained from a restricted model that excludes latitude as a predictor. This method of prodecure builds on an article published by @leocadio-miguel2017, which investigated the relationship between latitude and chronotype in the Brazilian population. The present study extends this work by utilizing a larger dataset, a more appropriate scale for measuring this phenomena [@leocadio-miguel2014], and more advanced statistical methods to test the hypothesis. The results will contribute to the ongoing debate on the latitude-chronotype relationship, offering new evidences into how environmental factors influence human circadian rhythms.

The thesis is structured as a [collection of articles ](https://en.wikipedia.org/wiki/Collection_of_articles). Chapters 2 through 4 comprise a series of essays and literature reviews related to the thesis topic. Chapter 5 presents the main investigation, including an article reporting the hypothesis test and discussing the central research question. Finally, Chapter 6 offers conclusions, limitations and directions for future research.

::: {.content-visible when-format="html"}
All the analyses in this document are reproducible and were conducted using the [R programming language](https://www.r-project.org/) [@rcoreteam] along with the [Quarto](https://quarto.org/) publishing system [@allaire].
:::

::: {.content-visible when-format="pdf"}
All analyses in this document are reproducible and were conducted using the [R programming language](https://www.r-project.org/) [@rcoreteam] along with the [Quarto](https://quarto.org/) publishing system [@allaire]. Since this thesis focuses heavily on data, it is best viewed online. To access the digital version and supplemental materials, please visit  [https://danielvartan.github.io/mastersthesis](https://danielvartan.github.io/mastersthesis/). The thesis code repository is available on GitHub at [https://github.com/danielvartan/mastersthesis)](https://github.com/danielvartan/mastersthesis). Finally, the research compendium can be accessed via [The Open Science Framework](https://osf.io/) at the following link:  [https://doi.org/10.17605/OSF.IO/YGKTS](https://doi.org/10.17605/OSF.IO/YGKTS).
:::
<!-- %:::% index end %:::% -->
