<!-- %:::% index begin %:::% -->
# Introduction

```{r}
#| label: setup
#| include: false

source(here::here("R/_setup.R"))
```

```{r}
#| label: status
#| echo: false
#| output: asis

rutils:::quarto_status(
  type = "drafting",
  of_what = "of this thesis",
  latex_parskip = "\\microskip"
  )
```

<!-- %:::% index pdf begin %:::% -->
You are currently viewing the preliminary `r ifelse(env_vars$format == "pdf", "print", "web")` version of this master's thesis.

This document follows the [collection of articles thesis](https://en.wikipedia.org/wiki/Collection_of_articles) format. This first chapter serves as an introduction to the thesis subject, providing also its justification and aims. The subsequent chapters consist of a series of essays and a article connected to the thesis, with the exception of the last one, which encompasses a general discussion and final remarks.

::: {.content-visible when-format="html"}
All analyses in this document are reproducible and were conducted using the [R programming language](https://www.r-project.org/) along with the [Quarto](https://quarto.org/) publishing system.
:::

::: {.content-visible when-format="pdf"}
All analyses in this document are reproducible and were conducted using the [R programming language](https://www.r-project.org/) along with the [Quarto](https://quarto.org/) publishing system. It's worth noting that this type of thesis is best suited for online viewing. To access the digital version and see the latest research updates, please visit [https://danielvartan.github.io/mastersthesis/](https://danielvartan.github.io/mastersthesis/).
:::

All of these projects are developed using secure, open-source tools and adhere to the best international standards. They are designed to ensure 100% reproducibility and are accompanied by extensive documentation.

## Aims

This thesis focuses on the expression of sleep and circadian phenotypes (chronotypes) with the aim of providing answers to the following questions:

1. How are the sleep-wake cycles and circadian phenotypes of the adult Brazilian population characterized?

1. Is latitude associated with the regulation of circadian rhythms in humans?

The basic hypothesis to be tested is that populations residing near the equator (latitude 0°) have, on average, a shorter/more morning-oriented circadian phenotype compared to populations living near the Earth's poles (H1) [@hut2013; @leocadio-miguel2014; @leocadio-miguel2017; @pittendrigh1991; @randler2008; @randler2017; @roenneberg2003].

The primary objectives (PO) are as follows:

```{=latex}
\microskip
```

A) Quantitatively describe the expression of sleep-wake cycles and circadian phenotypes of the Brazilian adult population at the end of the year 2017 (pre-pandemic).

A) Investigate and model the presence/absence of a significant association and effect between decimal degrees of latitude (independent variable (IV)) and circadian phenotypes (dependent variable (DV)) of the Brazilian population.

```{=latex}
\microskip
```

To achieve the primary objectives, the following secondary objectives (SO) have been outlined:

```{=latex}
\microskip
```

i) Conduct data cleaning, validation, and transformation processes on the obtained sample data.

i) Collect secondary data on geolocation and solarimetric models and cross-reference them with the primary data.

i) Develop algorithms for generating randomly sampled subsets adjusted to the proportions of the analyzed Brazilian regions, based on the latest Brazilian demographic census.

i) Develop algorithms and models to help with the processing of MCTQ data and to simulate the complexity of the entrainment phenomena.

i) Evaluate and discuss the presence/absence of significant differences in the values of the local time of the sleep corrected midpoint between sleep onset and sleep end on work-free days (MSF~sc~), MCTQ proxy for measuring the chronotype, based on decimal degrees of latitude (IV), while controlling for known covariates such as subjects' gender and age.

## Thesis structure

Three essays

The main project is the test of the latitude hypothesis, which serves as the primary investigation. It's important to note that all the preceding projects converge into this one. The first project focuses on validating the MCTQ translation used for data collection. The second project involves the development of data processing tools. The third project is responsible for the necessary data manipulation to prepare it for analysis. The fourth project aims to offer valuable insights and guidance for the upcoming tasks.
<!-- %:::% index end %:::% -->
