# Rule-based model of the 24h light/dark cycle entrainment phenomenon

```{r}
#| label: setup
#| echo: false
#| output: false

library(here)
source(here::here("R/quarto-setup.R"))
```

```{r}
#| echo: false
#| output: asis

quarto_status("drafting")
```

::: {.callout-note}
## Target

* [Journal of Open Source Software](https://joss.readthedocs.io/en/latest/submitting.html) ([IF 2022: 5.2/Exaly](https://exaly.com/journal/106890/journal-of-open-source-software/) | [B1/2017-2020](https://sucupira.capes.gov.br/sucupira/public/consultas/coleta/veiculoPublicacaoQualis/listaConsultaGeralPeriodicos.jsf)).
* [Journal of Statistical Software](https://www.jstatsoft.org/authors) ([IF 2022: 5.8/JCR](https://jcr.clarivate.com/jcr/) | [A1/2017-2020](https://sucupira.capes.gov.br/sucupira/public/consultas/coleta/veiculoPublicacaoQualis/listaConsultaGeralPeriodicos.jsf)).
:::

::: {.callout-note}
The following study was performed by __Daniel Vartanian (DV)__ and Camilo Rodrigues Neto (CR).

__DV__ contributed to the design, implementation of the software. __CR__ contributed as a science adviser and reviewer. __DV__ wrote the manuscript. All authors discussed the results and revised the final manuscript.

(_Future reference_): Vartanian, D, & Rodrigues Neto, C. (2024). {entrainment}: A rule-based model of the 24h light/dark cycle entrainment phenomenon. _Journal of Statistical Software_.
:::

## Study framework
