:::: {.panel-tabset}
### Age (years)

::: {#tbl-var-dist-stats-age}
```{r}
#| code-fold: true

data |>
  dplyr::pull(age) |>
  stats_summary(name = 'Age (years)')
```

Source: Created by the author.

Statistics for the `age` variable.
:::

::: {#fig-var-dist-hist-age}
```{r}
#| code-fold: true

data |>
  dplyr::pull(age) |>
  test_normality(name = 'Age (years)')
```

Source: Created by the author.

Histogram of the `age` variable with a kernel density
estimate, along with a quantile-quantile (Q-Q) plot between the
variable and the theoretical quantiles of the normal distribution.
:::

### Latitude (degrees)

::: {#tbl-var-dist-stats-latitude}
```{r}
#| code-fold: true

data |>
  dplyr::pull(latitude) |>
  stats_summary(name = 'Latitude (degrees)')
```

Source: Created by the author.

Statistics for the `latitude` variable.
:::

::: {#fig-var-dist-hist-latitude}
```{r}
#| code-fold: true

data |>
  dplyr::pull(latitude) |>
  test_normality(name = 'Latitude (degrees)')
```

Source: Created by the author.

Histogram of the `latitude` variable with a kernel density
estimate, along with a quantile-quantile (Q-Q) plot between the
variable and the theoretical quantiles of the normal distribution.
:::

### Longitude (degrees)

::: {#tbl-var-dist-stats-longitude}
```{r}
#| code-fold: true

data |>
  dplyr::pull(longitude) |>
  stats_summary(name = 'Longitude (degrees)')
```

Source: Created by the author.

Statistics for the `longitude` variable.
:::

::: {#fig-var-dist-hist-longitude}
```{r}
#| code-fold: true

data |>
  dplyr::pull(longitude) |>
  test_normality(name = 'Longitude (degrees)')
```

Source: Created by the author.

Histogram of the `longitude` variable with a kernel density
estimate, along with a quantile-quantile (Q-Q) plot between the
variable and the theoretical quantiles of the normal distribution.
:::

### MSF~sc~ (Chronotype proxy) (local time)

::: {#tbl-var-dist-stats-msf_sc}
```{r}
#| code-fold: true

data |>
  dplyr::pull(msf_sc) |>
  stats_summary(name = 'MSF~sc~ (Chronotype proxy) (local time)')
```

Source: Created by the author.

Statistics for the `msf_sc` variable.
:::

::: {#fig-var-dist-hist-msf_sc}
```{r}
#| code-fold: true

data |>
  dplyr::pull(msf_sc) |>
  test_normality(name = 'MSF~sc~ (Chronotype proxy) (local time)')
```

Source: Created by the author.

Histogram of the `msf_sc` variable with a kernel density
estimate, along with a quantile-quantile (Q-Q) plot between the
variable and the theoretical quantiles of the normal distribution.
:::
::::
