<!-- %:::% slides-results begin %:::% -->
---

::: {.notes}
Distribution of the local time of the sleep corrected midpoint between sleep onset and sleep end on work-free days (MSF~sc~), MCTQ proxy for measuring the chronotype. The categorical cut-offs follow a quantile approach going from extremely early ($0 |- 0.11$) to the extremely late ($0.88 - 1$).
:::

```{r}
#| include: false
#| message: false
#| warning: false

source(here::here("R/utils.R"))

utc_minus_3_states <- c(
  "Amapá", "Pará", "Maranhão", "Tocantins", "Piauí", "Ceará",
  "Rio Grande do Norte", "Paraíba", "Pernambuco", "Alagoas", "Sergipe",
  "Bahia", "Distrito Federal", "Goiás", "Minas Gerais", "Espírito Santo",
  "Rio de Janeiro", "São Paulo", "Paraná", "Santa Catarina",
  "Rio Grande do Sul"
)

data <- 
  targets::tar_read("geocoded_data", store = here::here("_targets")) |>
  dplyr::filter(state %in% utc_minus_3_states) |>
  dplyr::select(msf_sc, age, sex, state, latitude, longitude) |>
  tidyr::drop_na(msf_sc, age, sex, latitude)
```

```{r}
#| echo: false
#| message: false
#| warning: false

source(here::here("R/plot_chronotype.R"))

data |> 
  plot_chronotype(
    col = "msf_sc", 
    x_lab = "Frequency (%)",
    y_lab = latex2exp::TeX("Local time ($MSF_{sc}$)"),
    col_width = 0.8,
    col_border = 0.6,
    text_size = 12,
    chronotype_cuts = FALSE,
    legend_position = "right"
  )
```

---

::: {.notes}
Distribution of mean aggregates of the local time of the sleep corrected midpoint between sleep onset and sleep end on work-free days (MSF~sc~), MCTQ proxy for measuring the chronotype, in relation to latitude decimal degree intervals. Higher values of MSF~sc~ indicate a tendency toward a late chronotype. The red line represents a linear regression, and the shaded area indicates a pointwise 95% confidence interval.
:::

```{r}
#| echo: false
#| message: false
#| warning: false

source(here::here("R/plot_latitude_series.R"))

data |>
  dplyr::filter(age <= 50) |>
  plot_latitude_series(
    col = "msf_sc", 
    y_lab = latex2exp::TeX("$MSF_{sc} \\pm SEM$"), 
    line_width = 2, 
    point_size = 3,
    error_bar_width = 0.5, 
    error_bar_linewidth = 1, 
    error_bar = TRUE,
    text_size = 12
  )
```
<!-- %:::% slides-results end %:::% -->
