<!-- %:::% .common h1 begin %:::% -->
# Comparative Assessment of QualoCEP and Google Geocoding API Performance
<!-- %:::% .common h1 end %:::% -->

```{r}
#| label: setup
#| include: false

source(here::here("R", "_setup.R"))
```

## Overview

This is a simple workflow to test the [QualoCEP](https://www.qualocep.com/) geocoding against [Google Geocoding API](https://developers.google.com/maps/documentation/geocoding/overview).

This thesis uses both services for geocoding the data. QualoCEP is a paid service that provides a large dataset of Brazillian postal codes and addresses. The aim is to compare the results of QualoCEP's latitute e longitude values with Google's results for consistency.

::: {.callout-note}
To run this notebook you will need to have a Google Geocoding API key in your [`.Renviron`](https://bookdown.org/csgillespie/efficientR/set-up.html#:~:text=2.4.6%20The%20.Renviron%20file) file. Learn more about it [here](https://jessecambon.github.io/tidygeocoder/articles/geocoder_services.html).
:::

## Setting the enviroment

```{r}
#| eval: false

library(leaflet)
library(targets)
library(tidygeocoder)
```

```{r}
#| output: false

source(here::here("R", "get_brazil_fu.R"))
source(here::here("R", "get_qualocep_data.R"))
source(here::here("R", "render_brazil_address.R"))
source(here::here("R", "utils.R"))
```

```{r}
#| include: false

set.seed(2024)
```

## Importing QualOCep data

```{r}
#| output: false

qualocep_data <- get_qualocep_data()
```

## Random test 1

### QualOCep data

```{r}
sample_data <- dplyr::sample_n(qualocep_data, 1)

sample_data |> as.list() |> list_as_tibble()
```

```{r}
leaflet::leaflet() |> 
  leaflet::addTiles() |> 
  leaflet::addMarkers(
    lng = sample_data$longitude,
    lat = sample_data$latitude,
    popup = paste0(
      "Latitude: ", sample_data$latitude, "<br>",
      "Longitude: ", sample_data$longitude
    )
  )
```

### Google Geocoding API

```{r}
google_data <- 
  sample_data |>
  dplyr::mutate(
    address = render_brazil_address(
      street = sample_data$street,
      complement = sample_data$complement,
      neighborhood = sample_data$neighborhood,
      municipality = sample_data$municipality,
      state = sample_data$state,
      postal_code = sample_data$postal_code
    )
  ) |>
  tidygeocoder::geocode(
    address = address,
    method = "google"
  )

google_data |> as.list() |> list_as_tibble()
```

```{r}
leaflet::leaflet() |> 
  leaflet::addTiles() |> 
  leaflet::addMarkers(
    lng = google_data$longitude,
    lat = google_data$latitude,
    popup = paste0(
      "Latitude: ", google_data$latitude, "<br>",
      "Longitude: ", google_data$longitude
    )
  )
```

## Random test 2

### QualOCep data

```{r}
sample_data <- dplyr::sample_n(qualocep_data, 1)

sample_data |> as.list() |> list_as_tibble()
```

```{r}
leaflet::leaflet() |> 
  leaflet::addTiles() |> 
  leaflet::addMarkers(
    lng = sample_data$longitude,
    lat = sample_data$latitude,
    popup = paste0(
      "Latitude: ", sample_data$latitude, "<br>",
      "Longitude: ", sample_data$longitude
    )
  )
```

### Google Geocoding API

```{r}
google_data <- 
  sample_data |>
  dplyr::mutate(
    address = render_brazil_address(
      street = sample_data$street,
      complement = sample_data$complement,
      neighborhood = sample_data$neighborhood,
      municipality = sample_data$municipality,
      state = sample_data$state,
      postal_code = sample_data$postal_code
    )
  ) |>
  tidygeocoder::geocode(
    address = address,
    method = "google"
  )

google_data |> as.list() |> list_as_tibble()
```

```{r}
leaflet::leaflet() |> 
  leaflet::addTiles() |> 
  leaflet::addMarkers(
    lng = google_data$longitude,
    lat = google_data$latitude,
    popup = paste0(
      "Latitude: ", google_data$latitude, "<br>",
      "Longitude: ", google_data$longitude
    )
  )
```

## Random test 3

### QualOCep data

```{r}
sample_data <- dplyr::sample_n(qualocep_data, 1)

sample_data |> as.list() |> list_as_tibble()
```

```{r}
leaflet::leaflet() |> 
  leaflet::addTiles() |> 
  leaflet::addMarkers(
    lng = sample_data$longitude,
    lat = sample_data$latitude,
    popup = paste0(
      "Latitude: ", sample_data$latitude, "<br>",
      "Longitude: ", sample_data$longitude
    )
  )
```

### Google Geocoding API

```{r}
google_data <- 
  sample_data |>
  dplyr::mutate(
    address = render_brazil_address(
      street = sample_data$street,
      complement = sample_data$complement,
      neighborhood = sample_data$neighborhood,
      municipality = sample_data$municipality,
      state = sample_data$state,
      postal_code = sample_data$postal_code
    )
  ) |>
  tidygeocoder::geocode(
    address = address,
    method = "google"
  )

google_data |> as.list() |> list_as_tibble()
```

```{r}
leaflet::leaflet() |> 
  leaflet::addTiles() |> 
  leaflet::addMarkers(
    lng = google_data$longitude,
    lat = google_data$latitude,
    popup = paste0(
      "Latitude: ", google_data$latitude, "<br>",
      "Longitude: ", google_data$longitude
    )
  )
```
