<!-- %:::% .common h1 begin %:::% -->
# Power Analysis {#sec-sm-power-analysis}
<!-- %:::% .common h1 end %:::% -->

```{r}
#| label: setup
#| include: false

source(here::here("R", "_setup.R"))
```

## Overview

This document presents a power analysis for the Minimal Effect Size (MES) considered in the hypothesis test.

For an in-depth discussion of the thesis methods, see Supplementary Material [-@sec-sm-methods].

## Setting the Environment

```{r}
#| eval: false

library(pwrss)
```

## Hypothesis Test A

```{r}
#| include: false

pwr_analysis <- pwrss::pwrss.f.reg(
  f2 = 0.02, # Minimal Effect Size (MES)
  k = 8, # Number of predictors
  power = 0.99,
  alpha = 0.01
)
```

The results indicate that at least `{r} pwr_analysis$n |> groomr:::format_to_md_latex()` observations per variable are required to achieve a power of $0.99$ ($1 - \beta$) and a significance level ($\alpha$) of $0.01$ for the Test **A**. The dataset contains `{r} res_vars$ea_analysis_sample_n |> groomr:::format_to_md_latex()` observations, which exceeds this requirement.

```{r}
#| code-fold: false

pwr_analysis <- pwrss::pwrss.f.reg(
  f2 = 0.02, # Minimal Effect Size (MES)
  k = 8, # Number of predictors
  power = 0.99,
  alpha = 0.01
)
```

::: {#fig-sm-5-power-analysis-test-a}
```{r}
#| code-fold: true

pwrss::power.f.test(
  ncp = pwr_analysis$ncp,
  df1 = pwr_analysis$df1,
  df2 = pwr_analysis$df2,
  alpha = pwr_analysis$parms$alpha,
  plot = TRUE
)
```

[Source: Created by the author.]{.legend}

Visual representation of the power analysis for Test **A**.
:::

```{r}
#| include: false

quartor:::write_in_results_yml(
  list(
    hta_pwr_analysis = pwr_analysis
  )
)
```

## Hypothesis Test B

```{r}
#| include: false

pwr_analysis <- pwrss::pwrss.f.reg(
  f2 = 0.02, # Minimal Effect Size (MES)
  k = 5, # Number of predictors
  power = 0.99,
  alpha = 0.01
)
```

The results indicate that at least `{r} pwr_analysis$n |> groomr:::format_to_md_latex()` observations per variable are required to achieve a power of $0.99$ ($1 - \beta$) and a significance level ($\alpha$) of $0.01$ for the Test **B**. The dataset contains `{r} res_vars$ea_analysis_sample_n |> groomr:::format_to_md_latex()` observations, which exceeds this requirement.

```{r}
#| code-fold: false

pwr_analysis <- pwrss::pwrss.f.reg(
  f2 = 0.02, # Minimal Effect Size (MES)
  k = 5, # Number of predictors
  power = 0.99,
  alpha = 0.01
)
```

::: {#fig-sm-5-power-analysis-test-b}
```{r}
#| code-fold: true

pwrss::power.f.test(
  ncp = pwr_analysis$ncp,
  df1 = pwr_analysis$df1,
  df2 = pwr_analysis$df2,
  alpha = pwr_analysis$parms$alpha,
  plot = TRUE
)
```

[Source: Created by the author.]{.legend}

Visual representation of the power analysis for Test **B**.
:::

```{r}
#| include: false

quartor:::write_in_results_yml(
  list(
    htb_pwr_analysis = pwr_analysis
  )
)
```
