<!-- %:::% .common h1 begin %:::% -->
# Question and  hypothesis
<!-- %:::% .common h1 end %:::% -->

```{r}
#| label: setup
#| include: false

source(here::here("R", "_setup.R"))
options(scipen = 10, digits = 5)
```

```{r}
#| label: status
#| echo: false
#| output: asis

rutils:::quarto_status(
  type = "drafting",
  of_what = "of this thesis",
  latex_parskip = "\\microskip"
  )
```

```{r}
#| include: false

library(magrittr)
library(targets)
```

```{r}
#| include: false

source(here::here("R", "utils.R"))
```

```{r}
#| eval: false
#| output: false
#| code-fold: false

targets::tar_make(script = "../_targets.R")
```

```{r}
#| output: false

geocoded_data <- targets::tar_read(
  "geocoded_data", 
  store = here::here("_targets")
)
```

```{r}
#| output: false

weighted_data <- targets::tar_read(
  "weighted_data", 
  store = here::here("_targets")
)
```

## Overview

The aim of this document is to reinstate and detail the thesis object and objective, along with the methods and steps taken to build the models and test the hypothesis.

## Question

Every scientific work should be guided by a question. The question that guides this one is the following:

::: {.center-x}
**Is latitude associated with chronotype?**
:::

The primary objective is to model and test the hypothesis ingrained in the question, in the context of human circadian rhythms, by critically examining whether a significant association and effect size exist between latitude and circadian phenotypes in the Brazilian population.

## Hypothesis

To approach this question, I applied Popper’s hypothetico-deductive method, also known as the *method of conjecture and refutation* [@popper1979, p. 164]. The basic structure of this approach can be summarized as follows:

```{mermaid}
%%| label: fig-mermaid
%%| fig-cap: Simplified schema of Popper’s hypothetico-deductive method.
%%| fig-align: center

flowchart LR
  A(P1) --> B(TT)
  B --> C(EE)
  C --> D(P2)
```

> Here $\text{P}_1$, is the **problem** from which we start, $\text{TT}$ (the ‘tentative theory’) is the imaginative conjectural solution which we first reach, for example our first **tentative interpretation**. $\text{EE}$ (‘**error- elimination**’) consists of a severe critical examination of our conjecture, our tentative interpretation: it consists, for example, of the critical use of documentary evidence and, if we have at this early stage more than one conjecture at our disposal, it will also consist of a critical discussion and comparative evaluation of the competing conjectures. $\text{P}_2$ is the problem situation as it emerges from our first critical attempt to solve our problems. It leads up to our second attempt (**and so on**) [@popper1979, p. 164].

My tentative theory or main hypothesis, as stated in Chapter 1, was the following:

__Hypothesis__
: *Latitude is associated with chronotype distributions*, with populations closer to the equator exhibiting, on average, a shorter or more morning-oriented circadian phenotype compared to those residing near the poles.

This hypothesis traces back to the first discussions about chrnobiology (as a discipline) and can be seen on many studies, like: @hut2013; @leocadio-miguel2014; @leocadio-miguel2017; @pittendrigh1991; @randler2008; @randler2017; @roenneberg2003.

As a procedure method, an improved approach to Null Hypothesis Significance Testing ([NHST](https://en.wikipedia.org/wiki/Statistical_hypothesis_test)), rooted in the original Neyman-Pearson framework for data testing is employed [@neyman1928; @neyman1928a; @perezgonzalez2015] i the schema that follows.

$$
\begin{cases}
\text{H}_{0}: \text{Latitude is not associated with chronotype} \\
\text{H}_{a}: \text{Latitude is associated with chronotype}
\end{cases}
$$
