# Thesis outline {.unnumbered}

```{r}
#| label: setup
#| echo: false
#| output: false

library(here)
source(here::here("R/quarto-setup.R"))
```

```{r}
#| echo: false
#| output: asis

quarto_status("polishing")
```

[...]



All articles shown in the next chapters are target to specific science journals and will be submitted for publication soon.

While this thesis was being developed, the publication below were produced:

Viana-Mendes, J., Benedito-Silva, A. A., Andrade, M. A. M., Vartanian, D., Gonçalves, B. da S. B., Cipolla-Neto, J., & Pedrazzoli, M. (2023). Actigraphic characterization of sleep and circadian phenotypes of PER3 gene VNTR genotypes. _Chronobiology International_. https://doi.org/10.1080/07420528.2023.2256858

The author is also currently working with the development of the project below:

Sales, A. R. V., Vartanian, D., Andrade, M. A. M., Pedrazzoli, M. (2023). Associations between the duration and quality of sleep in third-trimester pregnant women and the duration of labor (Master's project). School of Arts, Sciences and Humanities, University of Sao Paulo, São Paulo.
