# Welcome {.unnumbered}

```{r}
#| label: setup
#| echo: false
#| output: false

library(here)
source(here::here("R/quarto-setup.R"))
```

```{r}
#| echo: false
#| output: false

# TODO: Move this function to {groomr}.

source(here::here("R/bbt_write_quarto_bib.R"))

bbt_write_quarto_bib(
  bib_file = here::here("references.json"), 
  dir = here::here(c("", "qmd")),
  pattern = "\\.qmd$"
  )
```

```{r}
#| echo: false
#| output: asis

quarto_status("drafting")
```

This is the web version of the master's thesis titled _Ecology of Sleep and Circadian Phenotypes of the Brazilian Population_. The printed version will soon be available at [teses.usp.br](https://teses.usp.br/).

The analyses contained in this thesis are 100% reproducible. They were performed using the [R programming language](https://www.r-project.org/) and the  [Quarto](https://quarto.org/) publishing system (an evolution of the [R Markdown](https://rmarkdown.rstudio.com/) system). Click on the GitHub icon in the menu to access the thesis code repository.

## Citation

For attribution, please cite this work as:

```
Vartanian, D. (2023). Ecology of sleep and circadian phenotypes of the 
Brazilian population [Master's thesis]. School of Arts, Sciences and Humanities,
University of Sao Paulo, São Paulo. https://danielvartan.github.io/mastersthesis/
```

BibTeX citation:

```
@mastersthesis{vartanian_2023,
  title = {Ecology of sleep and circadian phenotypes of the Brazilian population},
  author = {Daniel Vartanian},
  year = {2023},
  address = {São Paulo},
  school  = {University of Sao Paulo},
  langid = {portuguese},
  url = {https://danielvartan.github.io/mastersthesis/},
}
```
