# mastersthesis

<!-- quarto render --profile gfm -->

```{r}
#| label: setup
#| include: false

library(badger)

source(here::here("R", "_setup.R"))
```

<!-- badges: start -->
```{r}
#| echo: false
#| output: asis

cat(
  badger::badge_repostatus("active"),
  badger::badge_custom(
    "OSF DOI",
    "10.17605/OSF.IO/YGKTS",
    "1284C5",
    "https://doi.org/10.17605/OSF.IO/YGKTS"
  ),
  badger::badge_license(
    "MIT",
    "green",
    url = "https://choosealicense.com/licenses/mit/"
  ),
  badger::badge_license(
    "CC BY 4.0",
    "lightgrey",
    url = "https://creativecommons.org/licenses/by/4.0/"
  )
)
```
<!-- badges: end -->

## Overview

This repository contains the [Quarto book](https://quarto.org/docs/books/) of my Master's Thesis: *Is Latitude Associated with Chronotype?* The research compendium is accessible via [The Open Science Framework](https://osf.io/) by clicking [here](https://doi.org/10.17605/OSF.IO/YGKTS).

The assemble of this repository was inspired by Ben Marwick, Carl Boettiger & Lincoln Mullen's article [Packaging Data Analytical Work Reproducibly Using R (and Friends)](https://doi.org/10.1080/00031305.2017.1375986).

## How to Use

All analyses in this thesis are fully reproducible and were conducted using the [R programming language](https://www.r-project.org/) alongside the [Quarto](https://quarto.org/) publishing system. The [`renv`](https://rstudio.github.io/renv/) package was used to ensure that the R environment used can be restored (see `renv.lock`). All the computational notebooks can be found in the `qmd` directory.

It's important to note that some restrictions apply to the availability of the  main research data, which contain personal and sensitive information. As a result, this data cannot be publicly shared. To run the analyses, users must have an internet connection and request a set of access keys from the author (see *Keys* section).

If you don't feel comfortable with R, I strongly recommend checking Hadley Wickham and Garrett Grolemund's free and online book [R for Data Science](https://r4ds.hadley.nz/) and the Coursera course from John Hopkins University [Data Science: Foundations using R](https://www.coursera.org/specializations/data-science-foundations-r).

To reproduce the analyses do the following steps:

(@) Make sure that you have the latest R version and the following R packages installed in your machine:

```r
install.packages(
  c("cli", "here" ,"magrittr", "ragg", "renv", "stats", "stringr")
)
```

(@) Clone this repository.
(@) Open the R project in your IDE of choice.
(@) Run [`renv::restore()`](https://rstudio.github.io/renv//reference/restore.html) to install all software dependencies.
(@) Set all the keys provided by the author in the `.Renviron` file and the `_ssh` folder.
(@) Execute `targets::tar_make()` to run the entire pipeline, processing the raw data through to the validated and weighted data.
(@) Run any analysis present in the computational notebooks (`qmd` directory).

## Keys

To access the data and run the notebooks, you will need to set the following keys in a file named [`.Renviron`](https://bookdown.org/csgillespie/efficientR/set-up.html#:~:text=2.4.6%20The%20.Renviron%20file) located in the root directory of the project:

- `OSF_PAT`: Your [OSF](https://osf.io/) Personal Access Token (PAT). If you don't have one, go to the settings section of your OSF account and create a new token.
- `MASTERSTHESIS_PASSWORD`: The password for the project's RSA private key (32 bytes).

Additionally, you will need the following keys in the project's `_ssh` folder:

- `id_rsa`: The project's private RSA key (RSA 4096 bits (OpenSSL)).
- `id_rsa.pub`: The project's public RSA key.

These project's keys will be provided by the author of the thesis upon reasonable request. As reasonable requests, I consider those made by researchers with a legitimate interest in the thesis topic, whose projects are of high quality, ethically conducted, and who have obtained all necessary approvals for their research proposal.

## License

```{r}
#| echo: false
#| output: asis

cat(
  badger::badge_license(
    "MIT",
    "green",
    url = "https://choosealicense.com/licenses/mit/"
  ),
  badger::badge_license(
    "CC BY 4.0",
    "lightgrey",
    url = "https://creativecommons.org/licenses/by/4.0/"
  )
)
```

The code accompanying this thesis is distributed under the [MIT License](https://opensource.org/license/mit). All documents are released under the [Creative Commons Attribution 4.0 International Public License](https://creativecommons.org/licenses/by/4.0/).

The research data is subject to a private license and is not publicly available due to privacy and ethical considerations.

## How to Cite

To cite this work, please use the following format:

Vartanian, D. (`r env_vars[["submitted"]]`). *`r env_vars[["title"]] |> stringr::str_to_sentence()`* [`r env_vars[["academic-title"]]`'s `r env_vars[["type-of-work"]]`, `r env_vars[["university"]]`]. [https://doi.org/10.17605/OSF.IO/YGKTS](https://doi.org/10.17605/OSF.IO/YGKTS)

A BibTeX entry for LaTeX users is:

```
@mastersthesis{vartanian`r env_vars[["submitted"]]`,
  title = {`r env_vars[["title"]] |> stringr::str_to_sentence()`},
  author = {Daniel Vartanian},
  year = {`r env_vars[["year"]]`},
  address = {`r env_vars[["location"]]`},
  school  = {`r env_vars[["university"]]`},
  langid = {`r env_vars[["language"]]`},
  url = {https://doi.org/10.17605/OSF.IO/YGKTS},
  note = {`r env_vars[["version-note"]]`}
}
```

## Acknowledgments

```{r}
#| echo: false
#| output: asis

htmltools::tags$table(
  htmltools::tags$tr(
    htmltools::tags$td(
      htmltools::tags$br(),
      htmltools::tags$br(),
      htmltools::tags$p(
        htmltools::tags$a(
          htmltools::tags$img(
            src = "images/usp-logo.svg",
            width = 150
          ),
          href = paste0(
            "https://www.prpg.usp.br/pt-br/faca-pos-na-usp/",
            "programas-de-pos-graduacao/621-modelagem-de-sistemas-complexos"
          )
        ),
        align = "center"
      ),
      htmltools::tags$br(),
      width = "30%"
    ),
    htmltools::tags$td(
      htmltools::tags$p(
        "This thesis was developed in the Graduate Program in Modeling ",
        "Complex Systems (",
        htmltools::tags$a(
          "PPGSCX",
          href = paste0(
            "https://www.prpg.usp.br/pt-br/faca-pos-na-usp/",
            "programas-de-pos-graduacao/621-modelagem-de-sistemas-complexos"
          ),
          .noWS = "outside"
        ),
        ") at the University of São Paulo (",
        htmltools::tags$a(
          "USP",
          href = "https://www5.usp.br/",
          .noWS = "outside"
        ),
        ") under the supervision of",
        htmltools::tags$a(
          "Prof. Dr. Camilo Rodrigues Neto",
          href = "https://orcid.org/0000-0001-6783-6695",
          .noWS = "after"
        ),
        "."
      ),
      width = "70%"
    )
  )
) |>
  knitr::raw_html()

htmltools::tags$table(
  htmltools::tags$tr(
    htmltools::tags$td(
      htmltools::tags$br(),
      htmltools::tags$p(
        htmltools::tags$a(
          htmltools::tags$img(
            src = "images/capes-logo-original.svg",
            width = 120
          ),
          href = "https://www.gov.br/capes/"
        ),
        align = "center"
      ),
      htmltools::tags$br(),
      width = "30%"
    ),
    htmltools::tags$td(
      htmltools::tags$p(
        "This study was financed by the Coordenação de Aperfeiçoamento de ",
        "Pessoal de Nível Superior - Brazil (",
        htmltools::tags$a(
          "CAPES",
          href = "https://www.gov.br/capes/",
          .noWS = "outside"
        ),
        ") - Finance Code 001, Grant Number 88887.703720/2022-00."
      ),
      width = "70%"
    )
  )
) |>
  knitr::raw_html()
```
