---
title: "Text similarity"
description: |
  Draft notebook.
bibliography: ../references.bib
format:
  html:
    citations-hover: true
    code-copy: true
    code-overflow: wrap
    code-tools:
      source: https://github.com/danielvartan/mastersthesis
      toggle: false
      caption: none
    footnotes-hover: true
editor: source
highlight-style: arrow
lang: en
---

<!-- Go to <https://quarto.org/docs/guide/> for Quarto's documentation. -->

::: {.callout-note}
This document is just a draft.
:::

## First things first

```{r}
#| eval: false

library(checkmate)
library(cli)
library(curl)
library(here)
library(stringr)
library(text)
library(textreuse)
```

## Load and embed texts

Vartanian, D., & Pedrazzoli, M. (2017). Questionário de cronotipo: baseado no Munich ChronoType Questionnaire (MCTQ). https://giperbio.github.io/mctq-form/

Roenneberg, T., & Merrow, M. (2006). EUCLOCK: Portuguese MCTQ. http://web.archive.org/web/20141115175303/https://www.bioinfo.mpg.de/mctq/core_work_life/core/core.jsp?language=por_b

Reis, C., Madeira, S. G., Lopes, L. V., Paiva, T., & Roenneberg, T. (2020). Validation of the Portuguese variant of the Munich Chronotype Questionnaire (MCTQ-PT). Frontiers in Physiology, 11, 795. https://doi.org/10.3389/fphys.2020.00795

```{r}
field_form_text_file <- "data/field-form_core_main.txt"
euclock_text_file <- "data/euclock_core_main.txt"
mctq_pt_text_file <- "data/mctq-pt_core_main.txt"
```

```{r}
field_form_text_textreuse <- 
  textreuse::TextReuseTextDocument(
    file = field_form_text_file,
    meta = list(id = "field_form")
  )

euclock_text_textreuse <- 
  textreuse::TextReuseTextDocument(
    file = euclock_text_file,
    meta = list(id = "euclock")
  )

mctq_pt_text_textreuse <- 
  textreuse::TextReuseTextDocument(
    file = mctq_pt_text_file,
    meta = list(id = "mctq_pt")
  )
```

```{r}
#| output: false

if (!curl::has_internet()) {
  cli::cli_abort(paste0(
    "This code chunk needs an internet connection."
  ))
}

## See:
## https://huggingface.co/neuralmind/bert-base-portuguese-cased

text_embed <- function(file) {
  checkmate::assert_string(file)
  checkmate::assert_file_exists(file)
  
  file |>
    readLines() |>
    text::textEmbed(
      model = "neuralmind/bert-base-portuguese-cased",
      layers = - 2,
      dim_name = TRUE,
      aggregation_from_layers_to_tokens = "concatenate",
      aggregation_from_tokens_to_texts = "mean",
      aggregation_from_tokens_to_word_types = NULL,
      keep_token_embeddings = TRUE,
      tokens_select = NULL,
      tokens_deselect = NULL,
      decontextualize = FALSE,
      model_max_length = NULL,
      max_token_to_sentence = 4,
      tokenizer_parallelism = FALSE,
      device = "gpu",
      logging_level = "error"
    )
}

field_form_text_textembed <- text_embed(field_form_text_file)
euclock_text_textembed <- text_embed(euclock_text_file)
mctq_pt_text_textembed <- text_embed(mctq_pt_text_file)
```

## Text similarity

See @wang_2020 to learn more.

For a quick explanation, see: <https://youtu.be/e9U0QAFbfLI>.

<!-- See also: https://github.com/microsoft/semantic-kernel/blob/main/docs/EMBEDDINGS.md -->

```{r}
## See ?text::textSimilarity

text_distance <- function(x, y) {
  checkmate::assert_list(x, len = 2)
  checkmate::assert_list(y, len = 2)
  
  methods <- c(
    "binary", "cosine", "canberra", "euclidean", "manhattan", "maximum", 
    "minkowski", "pearson"
  )
  
  for (i in methods) {
    cli::cli_alert_info(paste0(
      "Method: {.strong {stringr::str_to_title(i)}}"
      ))
    
    test <- 
      text::textSimilarity(
        x$texts$texts, 
        y$texts$texts, 
        method = i, 
        center = TRUE, 
        scale = FALSE
      )
    
    cli::cli_bullets(c(">" = "Line by line"))
    print(test)
    
    cli::cli_bullets(c(">" = "Overall mean"))
    print(mean(test))
    
    cli::cat_line()
  }
}
```

```{r}
## See ?textreuse::jaccard_similarity

text_representation <- function(x, y) {
  checkmate::assert_class(x, "TextReuseTextDocument")
  checkmate::assert_class(y, "TextReuseTextDocument")
  
  cli::cli_alert_info(paste0("Method: {.strong Jaccard similarity}"))
  print(textreuse::jaccard_similarity(x, y))
  cli::cat_line()
  
  cli::cli_alert_info(paste0("Method: {.strong Jaccard bag similarity}"))
  print(textreuse::jaccard_bag_similarity(x, y))
  cli::cat_line()
}
```

### How similar is the _field form_ when compared to the _EUCLOCK form_?

* Text distance | Embedded (Semantic) test

```{r}
## See ?text::textSimilarity

text_distance(field_form_text_textembed, euclock_text_textembed)
```

* Text representation

```{r}
# The maximum value for the Jaccard bag similarity is 0.5.

text_representation(euclock_text_textreuse, field_form_text_textreuse)
```

### How similar is the _field form_ when compared to the _$\text{MCTQ}^{\text{PT}}$_?

* Text distance | Embedded (Semantic) test

```{r}
## See ?text::textSimilarity

text_distance(field_form_text_textembed, mctq_pt_text_textembed)
```

* Text representation

```{r}
# The maximum value for the Jaccard bag similarity is 0.5.

text_representation(mctq_pt_text_textreuse, field_form_text_textreuse)
```
